% !TEX root = ../../my-thesis.tex
\chapter{Implementation}
\label{sec:implementation}

This chapter details the complete software implementation of the GPU Cuckoo filter, translating the theoretical concepts from the previous chapter into a concrete, high-performance software artifact. The primary goal of the implementation is to provide a library that is not only fast but also robust, configurable, and scalable.

The chapter begins by outlining the fundamental design of the data structure itself, covering its compile-time configuration options, its memory layout optimized for parallel hardware, and its comprehensive public-facing API. Following this, the chapter dives into the core of the implementation: the parallel algorithms for insertion, lookup, and deletion. Each operation is presented as a pseudocode kernel, with detailed explanations of the logic used to handle concurrency, manage the eviction process, and ensure correctness.

Beyond the baseline algorithms, some advanced optimization techniques are explored. These include a sorted insertion strategy to improve memory locality for large filters, an alternative eviction method designed to reduce the amount of necessary evictions at high load factors, the option to load more fingerprints at once when using supported hardware and the usage of bucket placement policies other than the original XOR-based partial-key cuckoo hashing.

Finally, two major extensions are presented that elevate the filter from a single-GPU data structure to a system-level component. An IPC wrapper is detailed, which enables zero-copy sharing of the filter between multiple processes on the same machine. To address datasets larger than a single GPU's memory, the following section then describes a multi-GPU implementation that partitions the workload and data across multiple devices, using high-speed interconnects for efficient communication when possible.

\input{content/03-implementation/challenges}
\input{content/03-implementation/data-structure-design}
\input{content/03-implementation/parallel-algorithms}
\input{content/03-implementation/optimization-techniques}
\input{content/03-implementation/ipc-wrapper}
\input{content/03-implementation/multi-gpu}
\input{content/03-implementation/testing-verification}
